Supporting Information for ``Robust Adaptation to Multi-Scale Climate Variability''

### 2019-03-29 report-earths-future/si_template_2019.tex
incorporate scott suggestions

As described in Section XXX, we use a synthetic ENSO time series, specifically a NINO3 time series, from \cite{Ramesh:2016hf} to represent low-frequency climate variability.
Figure XXX shows the wavelet spectrum of this time series, calculated using the WaveletComp package \cite{Roesch:wlBQQoIs}.
The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of ENSO although interactions with other modes of LFV found in some studies \cite{Jin:1994wq} is dampened in this model.
There are, however, modes of LFV between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.

@@

As described in {sec:methods}, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
The full model, including priors, for the stationary log-normal model is given by {eq:log-normal-stationary}:
\begin{align}\label{eq:log-normal-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \sigma) \\
    \mu &\sim \mathcal{N} \qty(7, 1.5) \\
    \sigma &\sim \mathcal{N}^+ \qty(1, 1)
  \end{split}
\end{align}
where $\mathcal{N}$ denotes the normal distribution and $\mathcal{N}^+$ denotes a half-normal distribution.
The full model, including priors, for the trend log-normal model is given by {eq:log-normal-trend}:
\begin{align}\label{eq:log-normal-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \xi \mu) \\
    \mu_0 & \sim \mathcal{N} \qty(7, 1.5) \\
    \beta_\mu & \sim \mathcal{N} \qty(0, 0.1) \\
    \log \xi & \sim \mathcal{N} \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.

@@

In the main text we sampled climate risk by considering dependence on a single climate index, as shown in {eq:nino3}.
Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is given by
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
Parameters for the three experiments are given in {tab:markov-stationary}.
Results for these experiments are shown in {fig:lfv-markov-bias-variance,fig:lfv-secular-markov-bias-variance,fig:secular-only-markov-bias-variance}.

@@

\begin{figure}
      \includegraphics[width=\textwidth]{enso_wavelet.png}
      \caption{
        Wavelet analysis of the synthetic annual NINO3 time series described in Section XXX.
        (L): wavelet power spectrum.
        (R): global (average) power spectrum.
        Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
      }\label{fig:enso-ts}
    \end{figure}

    \begin{table}[ht]
      \centering
      \begin{tabular}{llll}
        \toprule
        Parameter & LFV Only & Secular Only & Secular + LFV \\
        \midrule
        $\mu_0$             & 6     & 6.5   & 6 \\
        $\gamma$            & 0     & 0.015 & 0.015\\
        $\beta$             & 0.5   & 0     & 0.5\\
        $\xi$               & 0.1   & 0.1   & 0.1\\
        $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
        Threshold           & 3000  & 3000  & 3000 \\
        \bottomrule
      \end{tabular}
      \caption{
        Parameters of {eq:nino3} for the LFV only, secular only, and secular plus LFV scenarios.
      }\label{tab:nino-stationary}
    \end{table}

    \begin{table}[ht]
      \centering
      \begin{tabular}{llll}
        \toprule
        Parameter & LFV Only & Secular Only & Secular + LFV \\
        \midrule
        $\mu_1$             & 6.75  & 6.5   & 6.75\\
        $\mu_2$             & 6     & 6.5   & 6\\
        $\gamma_1$          & 0     & 0.015 & 0.015\\
        $\gamma_2$          & 0     & 0.015 & 0\\
        $\xi$               & 0.1   & 0.1   & 0.1\\
        $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
        Threshold           & 3000  & 3000  & 3000 \\
        \bottomrule
      \end{tabular}
      \caption{
        Parameters of the two-state Markov chain model for the LFV only, secular only, and secular plus LFV experiments.
      }\label{tab:markov-stationary}
    \end{table}
    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
      \caption{
        LFV only: as {fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
      }\label{fig:lfv-markov-bias-variance}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{secular-only-markov-bias-variance.pdf}
      \caption{
        Secular change: as {fig:secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
      }\label{fig:secular-only-markov-bias-variance}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
      \caption{
        LFV plus secular change: as {fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
      }\label{fig:lfv-secular-markov-bias-variance}
    \end{figure}
\end{document}

>>>>

As described in Section 2.1, we use a synthetic ENSO time series, specifically a NINO3 time series, from \cite{Ramesh:2016hf} to represent low-frequency climate variability.
Figure S1 shows the wavelet spectrum of this time series, calculated using the WaveletComp package \cite{Roesch:wlBQQoIs}.
The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of ENSO although interactions with other modes of LFV found in some studies \cite{Jin:1994wq} is dampened in this model.
There are, however, modes of LFV between 16 and 128 years which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.

@@

As described in Section 2, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
The full model, including priors, for the stationary log-normal model is given by equation \eqref{log-normal-stationary}:
\begin{align}\label{eq:log-normal-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \sigma) \\
    \mu &\sim \mathcal{N} \qty(7, 1.5) \\
    \sigma &\sim \mathcal{N}^+ \qty(1, 1)
  \end{split}
\end{align}
where $\mathcal{N}$ denotes the normal distribution and $\mathcal{N}^+$ denotes a half-normal distribution.
The full model, including priors, for the trend log-normal model is given by equation \eqref{eq:log-normal-trend}:
\begin{align}\label{eq:log-normal-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \xi \mu) \\
    \mu_0 & \sim \mathcal{N} \qty(7, 1.5) \\
    \beta_\mu & \sim \mathcal{N} \qty(0, 0.1) \\
    \log \xi & \sim \mathcal{N} \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.

@@

In the main text we sampled climate risk by considering dependence on a single climate index, as shown in equation 5.
Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is given by
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
Parameters for the three experiments are given in Table \ref{tab:markov-stationary}.
Results for these experiments are shown in Figure \ref{fig:lfv-markov-bias-variance}, \ref{fig:lfv-secular-markov-bias-variance}, and \ref{fig:secular-only-markov-bias-variance}.

@@

\begin{figure}
      \includegraphics[width=\textwidth]{enso_wavelet.png}
      \caption{
        Wavelet analysis of the synthetic annual NINO3 time series described in Section 2.1.
        (L): wavelet power spectrum.
        (R): global (average) power spectrum.
        Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
      }\label{fig:enso-ts}
    \end{figure}

    \begin{table}[ht]
      \centering
      \begin{tabular}{llll}
        \toprule
        Parameter & LFV Only & Secular Only & Secular + LFV \\
        \midrule
        $\mu_0$             & 6     & 6.5   & 6 \\
        $\gamma$            & 0     & 0.015 & 0.015\\
        $\beta$             & 0.5   & 0     & 0.5\\
        $\xi$               & 0.1   & 0.1   & 0.1\\
        $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
        Threshold           & 3000  & 3000  & 3000 \\
        \bottomrule
      \end{tabular}
      \caption{
        Parameters of equation (5) for the LFV only, secular only, and secular plus LFV scenarios.
      }\label{tab:nino-stationary}
    \end{table}

    \begin{table}[ht]
      \centering
      \begin{tabular}{llll}
        \toprule
        Parameter & LFV Only & Secular Only & Secular + LFV \\
        \midrule
        $\mu_1$             & 6.75  & 6.5   & 6.75\\
        $\mu_2$             & 6     & 6.5   & 6\\
        $\gamma_1$          & 0     & 0.015 & 0.015\\
        $\gamma_2$          & 0     & 0.015 & 0\\
        $\xi$               & 0.1   & 0.1   & 0.1\\
        $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
        Threshold           & 3000  & 3000  & 3000 \\
        \bottomrule
      \end{tabular}
      \caption{
        Parameters of the two-state Markov chain model for the LFV only, secular only, and secular plus LFV experiments.
      }\label{tab:markov-stationary}
    \end{table}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{secular-only-markov-bias-variance.pdf}
      \caption{
        Secular change: as Figure 5 for sequences generated with the two-state Markov chain model.
      }\label{fig:secular-only-markov-bias-variance}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
      \caption{
        LFV only: as Figure 6 for sequences generated with the two-state Markov chain model.
      }\label{fig:lfv-markov-bias-variance}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
      \caption{
        LFV plus secular change: as Figure 7 for sequences generated with the two-state Markov chain model.
      }\label{fig:lfv-secular-markov-bias-variance}
    \end{figure}
\end{document}
